Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-05-03 Paper/paper.tex
Update text
This is from Gabor

We assume a flat cosmology with density parameters
$\left(\Omega_\mathrm{m},\Omega_\Lambda\right)=\left(0.3,0.7\right)$
and a Hubble constant $H_0=70$\,km\,s$^{-1}$\,Mpc$^{-1}$. Magnitudes
are reported in the AB system \citep{1983ApJ...266..713O}, and
observed magnitudes are point spread function (PSF) magnitudes
\citep{2002AJ....123..485S} corrected for Galactic extinction
\citep{1998ApJ...500..525S} unless otherwise noted.

@@

\begin{table*}
  \caption{Data sets analysed in this work.  We restrict the SDSS DR7
    data set \citep{2010AJ....139.2360S} to $z<4.7$.  The
    \citet{2013ApJ...768..105M} data set is restricted to faint
    ($M_{1450}>-26.73$) quasars to avoid overlap with the
    \citet{2016ApJ...829...33Y} sample.  Most AGN in the
    \citet{2015AA...578A..83G} sample do not have spectroscopic
    redshifts.  We therefore only use this sample for secondary
    analysis in Section~\ref{sec:global} and Appendix~\ref{sec:conv}.
    Similarly, the quasars reported by \citet{2011Natur.474..616M},
    \citet{2015ApJ...801L..11V}, and \citet{2018Natur.553..473B} have
    only approximate completeness estimates.  We only use these
    quasars in our secondary analysis in Section~\ref{sec:global}.}
  \label{tab:samples}
  \begin{tabular}{ccllrS}
    \hline
    & $z$ & Survey & Reference & Number & {Area} \\
    & & & & of quasars & {(deg$^2$)} \\
    \hline
    1. & 0.1--2.2 & SDSS DR7 & \citet{2010AJ....139.2360S} & 48664 & 6248.0 \\
    2. & 0.4--2.6 & 2SLAQ SGP & \citet{2009MNRAS.392...19C} & 3663 & 64.2 \\
    3. &  0.4--2.6 & --- NGP & \citet{2009MNRAS.392...19C} & 8153 & 127.7 \\
    4. & 2.2--3.5 & BOSS DR9 & \citet{2013ApJ...773...14R} & 23301 & 2236.0 \\
    5. & 3.7--4.7 & SDSS DR7 & \citet{2010AJ....139.2360S} & 1785 & 6248.0 \\
    6. & 4.7--5.4 & SDSS+WISE & \citet{2016ApJ...829...33Y} & 99 & 14555.0 \\
    7. & 4.7--5.5 & SDSS DR7 & \citet{2013ApJ...768..105M} & 103 & 6248.0 \\
    8. & 4.7--5.4 & --- Stripe 82 & \citet{2013ApJ...768..105M} & 59 & 235.0 \\
    9. &3.7--4.8 & NDWFS & \citet{2011ApJ...728L..26G} & 12 & 1.71 \\
    10. &3.8--5.1 & DLS & \citet{2011ApJ...728L..26G} & 12 & 2.05 \\
    11. &4.1--6.3 & CANDELS GOODS-S & \citet{2015AA...578A..83G} & 19 & 0.047 \\
    12. & 5.8--6.4 & SDSS Main & \citet{2016ApJ...833..222J} & 24 & 11240.0 \\
    13. & 5.9--6.1 & --- Overlap & \citet{2016ApJ...833..222J} & 10 & 4223.0 \\
    14. & 5.7--6.1 & --- Stripe 82 & \citet{2016ApJ...833..222J} & 13 & 277.0 \\
    15. & 6.0 & CFHQS Deep & \citet{2010AJ....139..906W} & 1 & 4.47 \\
    16. & 5.9--6.4 & --- Very Wide & \citet{2010AJ....139..906W} & 16 & 494.0 \\
    17. & 6.0--6.2 & Subaru High-$z$ Quasar & \citet{2015ApJ...798...28K} & 2 & 6.5 \\
    18. & 6.5--7.4 & UKIDSS & \citet{2011Natur.474..616M} & 1 & 3370.0 \\
    19. & 6.5--7.4 & UKIDSS & \citet{2015ApJ...801L..11V} & 1 & 3370.0 \\
    20. & 6.5--7.4 & ALLWISE+UKIDSS+DECaLS & \citet{2018Natur.553..473B} & 1 & 2500.0 \\
    \hline
  \end{tabular}\\
\end{table*}

@@

We started with compiling the samples of recent rest-frame UV-optical
quasar surveys. The restriction to UV-optical surveys was mainly
driven by our science goal to characterise the UV luminosity function
of Type~1 quasars. X-ray-selected samples are less suited for this
purpose due to spectroscopic incompleteness and the $\sim 0.4$\,dex
scatter in the conversion from X-ray to UV luminosity
\citep{2010A&A...512A..34L, 2015MNRAS.453.1946G, 2016ApJ...819..154L}
that contributes significantly to the error budget in the UV
luminosity function of X-ray-selected samples unless rest-frame UV
photometry is incorporated \citep{2015AA...578A..83G}.  The individual
surveys and their main characteristics are listed in
Table~\ref{tab:samples}.  Figure~\ref{fig:qsos} presents a redshift
histogram of the contributing surveys.

@@

At $4.7\le z<5.5$ we combined several recent surveys, accounting for
sample overlap and updated selection functions. At the bright end of
the QLF we used the 99 quasars from the SDSS+WISE survey
\citep{2016ApJ...829...33Y} that have $M_{1450}<-26.73$ in our adopted
cosmology.  For these 99 quasars selected from 14,555\,deg$^2$ we
adopted the \citet{2016ApJ...829...33Y} selection function.  The
\citet{2016ApJ...829...33Y} sample partially overlaps with the SDSS
DR7 sample from \citet{2013ApJ...768..105M}, so to avoid
double-counting quasars we used the latter sample only at $M_{1450}\ge
-26.73$, yielding 104 additional $4.7\le z<5.5$ quasars selected in
6248\,deg$^2$. We used the $z\sim 5$ SDSS DR7 selection function from
\citet{2013ApJ...768..105M} that supersedes the one from
\citet{2006AJ....131.2766R} due to improved $K$ corrections and IGM
parameterization.  To these two bright-end samples we added the
faint-end sample from the \citet{2013ApJ...768..105M} SDSS Stripe~82
survey (62 uniformly selected $4.7\le z<5.5$ quasars in 235\,deg$^2$)
and two $4.7\le z<5.5$ quasars from \citet{2011ApJ...728L..26G},
adopting the respective selection functions.  We did not consider the
limit on the $z\sim 5$ QLF by \citet{2012ApJ...756..160I} due to
systematic errors in their selection
function\footnote{\citet{2012ApJ...756..160I} underestimated the
  dispersion in rest-frame UV quasar colours with respect to SDSS at
  all redshifts (their Figure~4). Contrary to their claim, photometric
  errors have a small effect on the colour distribution of SDSS
  quasars given the statistical errors of $<0.03$\,mag in $gri$ for 90
  per cent of the SDSS DR7 bright quasar sample ($i<19.1$) and a
  relative calibration error of $\sim 1$ per cent
  \citep{2008ApJ...674.1217P}.}.

@@

For a joint fit of the QLF it is necessary to homogenise the different
survey samples in absolute magnitude, and to convert their selection
functions to the same absolute magnitude system. For the analysis of
the quasar UV emissivity and to be consistent with published work at
$z>3$ we chose to convert all samples and selection functions to the
absolute AB magnitude at a rest frame wavelength $\lambda=1450$\,\AA
\begin{equation}\label{eq:absmag}
  M_{1450}\left(z\right) = m-5\log{\left(\frac{d_L\left(z\right)}
    {\mathrm{Mpc}}\right)}-25-K_{m,1450}\left(z\right),
\end{equation}
with the luminosity distance $d_L\left(z\right)$ to a quasar at
redshift $z$, the apparent magnitude $m$ in a filter used in the
survey, and the bandpass correction $K_{m,1450}\left(z\right)$
\citep{1956AJ.....61...97H, 1968ApJ...154...21O, 2000A&A...353..861W,
  2002astro.ph.10394H}. For the bandpass correction we used a
combination of the \citet{2015MNRAS.449.4204L} quasar SED at
$\lambda<2500$\,\AA, and the \citet{2001AJ....122..549V} quasar
composite spectrum at longer wavelengths to cover the lowest
redshifts.  The samples from SDSS and BOSS are defined in the SDSS $i$
band, while 2SLAQ is defined in the $g$ band. At $z>4.7$ we adopted
the SDSS $z$ band magnitude (in the following denoted $z_\mathrm{AB}$)
for SDSS DR7 quasars to avoid additional corrections due to the
Ly$\alpha$ forest. Figure~\ref{fig:kcorr} shows our bandpass
corrections for SDSS, BOSS and 2SLAQ as a function of redshift. We
ignored the luminosity dependence of the bandpass correction due to
the known anticorrelation of emission line equivalent width and
luminosity \citep{1977ApJ...214..679B}.  While the
\citet{2015MNRAS.449.4204L} SED is for luminous ($M_{1450}\simeq
-27.2$) quasars, UV composite spectra including fainter quasars
\citep{2002ApJ...565..773T, 2012ApJ...752..162S, 2014ApJ...794...75S}
give similar values, such that our bandpass corrections remain
applicable at $M_{1450}\la -24$. Empirical luminosity-dependent
bandpass corrections show a $\la 0.2$\,mag variation over $\sim 4$
orders of absolute magnitude depending on redshift and the filter, and
with a $\sim 0.2$\,mag intrinsic scatter due to individual SED
variations \citep{2013ApJ...773...14R, 2013ApJ...768..105M,
  2013A&A...551A..29P}.

@@

For 2SLAQ we corrected for magnitude-dependent spectroscopic coverage
in the two survey areas \citep[$f_\mathrm{c,NGP}\left(g\right)$ and
  $f_\mathrm{c,SGP}\left(g\right)$; Figure~4
  in][]{2009MNRAS.392...19C} and spectroscopic redshift success
\citep[$f_\mathrm{s,2SLAQ}\left(g\right)$; Figure~6b
  in][]{2009MNRAS.392...19C} by multiplying them with the photometric
selection function, resulting in two area-specific 2SLAQ selection
functions
$f_\mathrm{NGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,NGP}f_\mathrm{s,2SLAQ}$
and
$f_\mathrm{SGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,SGP}f_\mathrm{s,2SLAQ}$
that are relevant for the QLF.  The $z<4.7$ SDSS photometric selection
function was modified to include known imaging incompleteness to
$f_{\mathrm{SDSS},z<4.7}=0.95f_{\mathrm{p,SDSS},z<4.7}$
\citep{2006AJ....131.2766R}. The BOSS colour-selected sample contains
quasars with $f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}\ge 0.85$
\citep{2013ApJ...773...14R}, and we adopted
$f_\mathrm{BOSS}=\overline{f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}}f_\mathrm{p,BOSS}=0.962f_\mathrm{p,BOSS}$. \citet{2010ApJ...710.1498G}
presented two area-specific photometric selection functions due to
different filters employed, and more follow-up spectroscopy was
reported in \citet{2011ApJ...728L..26G}. We accounted for remaining
spectroscopic incompleteness at $R>23$, yielding the final selection
functions $f_\mathrm{NDWFS}$ and $f_\mathrm{DLS}$. The updated $z\sim
5$ SDSS photometric selection function \citep{2013ApJ...768..105M} was
modified to include imaging and spectroscopic incompleteness, yielding
$f_{\mathrm{SDSS},z\sim 5}=0.95^2f_{\mathrm{p,SDSS},z\sim 5}$. In the
deeper $z\sim 5$ SDSS Stripe~82 survey the spectroscopic
incompleteness is larger and magnitude-dependent \citep[Figure~14
  in][]{2013ApJ...768..105M}, resulting in $f_{\mathrm{S82},z\sim
  5}=0.95f_{\mathrm{s,S82},z\sim
  5}\left(i\right)f_{\mathrm{p,S82},z\sim 5}$. Likewise, imaging and
magnitude-dependent spectroscopic incompleteness was factored into the
\citet{2016ApJ...829...33Y} photometric selection function (their
Figures~5 and 7), resulting in
$f_\mathrm{SDSS+WISE}=0.97f_\mathrm{s,SDSS+WISE}\left(z_\mathrm{AB}\right)f_\mathrm{p,SDSS+WISE}$.
Finally, we obtained a rough estimate of the
\citet{2015AA...578A..83G} selection function by comparing the
corrected and observed QLFs, i.e.\ taking
$f_\mathrm{GOODS-S}=\phi_\mathrm{obs}/\phi_\mathrm{corr}$ (see their
Table~3).

@@

Given quasar luminosities and estimates for survey completeness, we
are in a position to estimate the UV luminosity function of AGN.  We
begin our analysis by first deriving luminosity functions in bins of
redshift and magnitude.  In this discussion all magnitudes are
absolute magnitudes at rest-frame 1450~{\AA}.
% Put that sentence about magnitudes in the introduction.

@@

In a magnitude bin $[M_\mathrm{min}, M_\mathrm{max})$, and redshift
bin $[z_\mathrm{min}, z_\mathrm{max})$, we define the luminosity
function as \citep{2000MNRAS.311..433P}
\begin{equation}
  \phi \equiv \frac{N}{V_\mathrm{bin}},
\end{equation}
where $N$ is the number of quasars with magnitude
$M_\mathrm{min}\leq M<M_\mathrm{max}$ and redshift
$z_\mathrm{min}\leq z<z_\mathrm{max}$, and
\begin{equation}
  V_\mathrm{bin} = \int_{M_\mathrm{min}}^{M_\mathrm{max}}dM
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}dz\, f(M, z)\,\frac{dV}{dz},
  \label{eqn:vi}
\end{equation}
is the effective volume of the bin.  Here $f(M,z)$ is the quasar
selection probability, which includes incompleteness.  Inclusion of
the selection probability in Equation~(\ref{eqn:vi}) accounts for
what are sometimes called ``incomplete bins''
\citep{2006AJ....131.2766R}.  The comoving volume element $dV/dz$ is
given by
\begin{equation}
  \frac{dV}{dz}=\frac{dV}{dz\,d\Omega}\cdot A\cdot\frac{4\pi}{41253},
\end{equation}
where $A$ is the survey area in deg$^2$, and 
\begin{equation}
  \frac{dV}{dz\,d\Omega}=\frac{c}{H_0}\frac{d_L(z)^2}
       {(1+z)^2\sqrt{\Omega_m(1+z)^3+\Omega_\Lambda}},
  \label{eqn:dvdzdo}
\end{equation}
denotes the comoving volume element per unit solid angle
\citep{1999astro.ph..5116H}.  The luminosity distance $d_L$ is given
by
\begin{equation}
  d_L(z)=(1+z)\frac{c}{H_0}\int_0^z\frac{dz}
  {\sqrt{\Omega_m(1+z)^3+\Omega_\Lambda}}.
  \label{eqn:dl}
\end{equation}
Equations~(\ref{eqn:dvdzdo}) and (\ref{eqn:dl}) assume a flat Universe
($\Omega_k=0$).  The luminosity function $\phi$ has units of
$\mathrm{cMpc}^{-3}\mathrm{mag}^{-1}$.  We evaluate the double
integral in Equation~(\ref{eqn:vi}) by the Euler method, i.e., by
simply summing over the ``tiles'' of the selection function map,
without interpolating between the redshift and luminosity values of
neighbouring tiles.  This may result in $V_i=0$ for some quasars, in
which case we remove them from our analysis.

In each bin, we estimate the uncertainty in the luminosity function by
assuming Poisson statistics \citep{1986ApJ...303..336G} for the number
of quasars in each bin.  The resultant binned luminosity function
estimates are shown by the points in Figure~\ref{fig:mosaic}.

As seen in Figure~\ref{fig:mosaic}, the distribution of luminosity
function values in each redshift bin are suggestive of a double power
law form.  We will fit such a form below.  However, we also see in
several redshift bins of Figure~\ref{fig:mosaic} that the luminosity
function seems to show a decline at the faint end.  This is clearly
seen, for example, in the 2SLAQ sample at $z<2$ and in the SDSS quasar
sample at $z<1.8$ and $z\sim 4$.  We take this to be a systematic
error.  The decrease in the luminosity function is an indication of
further low-level incompleteness in the quasar surveys that has not
been captured by the reported completeness values.  We choose to
discard such bins from our analysis.  The discarded luminosity bins
are shown in Figure~\ref{fig:mosaic} by open circles.

>>>>

We assume a flat cosmology with density parameters
$\left(\Omega_\mathrm{m},\Omega_\Lambda\right)=\left(0.3,0.7\right)$
and a Hubble constant $H_0=70$\,km\,s$^{-1}$\,Mpc$^{-1}$. Comoving
distances are given explicitly in comoving Mpc (cMpc). Magnitudes
are reported in the AB system \citep{1983ApJ...266..713O}, and
observed magnitudes are point spread function (PSF) magnitudes
\citep{2002AJ....123..485S} corrected for Galactic extinction
\citep{1998ApJ...500..525S} unless otherwise noted.
Our homogenised sample (Section~\ref{sec:sample}) uses absolute
monochromatic AB magnitudes at a rest frame wavelength of 1450\,\AA.

@@

\begin{table*}
  \caption{Quasar samples analysed in this work.}
  \label{tab:samples}
  \begin{tabular}{lcllrS}
    \hline
    Sample& $z$ range$^a$& Survey & Reference & Number & {Area} \\
    & & & & of quasars & {(deg$^2$)} \\
    \hline
    1 & 0.1--2.2 & SDSS DR7 & \citet{2010AJ....139.2360S} & 48664 & 6248.0 \\
    2 & 0.4--2.6 & 2SLAQ SGP & \citet{2009MNRAS.392...19C} & 3663 & 64.2 \\
    3 & 0.4--2.6 & --- NGP & \citet{2009MNRAS.392...19C} & 8153 & 127.7 \\
    4 & 2.2--3.5 & BOSS DR9 & \citet{2013ApJ...773...14R} & 23301 & 2236.0 \\
    5 & 3.7--4.7 & SDSS DR7 & \citet{2010AJ....139.2360S} & 1785 & 6248.0 \\
    6 & 3.6--5.2 & NDWFS & \citet{2011ApJ...728L..26G} & 12 & 1.71 \\
    7 & 3.8--5.3 & DLS & \citet{2011ApJ...728L..26G} & 12 & 2.05 \\
    8 & 4.7--5.4 & SDSS+WISE & \citet{2016ApJ...829...33Y} & 99 & 14555.0 \\
    9$^b$ & 4.7--5.5 & SDSS DR7 & \citet{2013ApJ...768..105M} & 103 & 6248.0 \\
    10 & 4.7--5.5 & --- Stripe 82 & \citet{2013ApJ...768..105M} & 59 & 235.0 \\
    11 & 5.7--6.5 & SDSS Main & \citet{2016ApJ...833..222J} & 24 & 11240.0 \\
    12 & 5.7--6.5 & --- Overlap & \citet{2016ApJ...833..222J} & 10 & 4223.0 \\
    13 & 5.7--6.5 & --- Stripe 82 & \citet{2016ApJ...833..222J} & 13 & 277.0 \\
    14 & 5.8--6.6 & CFHQS Deep & \citet{2010AJ....139..906W} & 1 & 4.47 \\
    15 & 5.8--6.6 & --- Very Wide & \citet{2010AJ....139..906W} & 16 & 494.0 \\
    16 & 5.8--6.5 & Subaru High-$z$ Quasar & \citet{2015ApJ...798...28K} & 2 & 6.5 \\
    17$^c$ &4.0--6.5 & CANDELS GOODS-S & \citet{2015AA...578A..83G} & 19 & 0.047 \\
    18$^d$ & 6.5--7.4 & UKIDSS & \citet{2011Natur.474..616M} & 1 & 3370.0 \\
    19$^d$ & 6.5--7.4 & UKIDSS & \citet{2015ApJ...801L..11V} & 1 & 3370.0 \\
    20$^d$ & 6.5--7.4 & ALLWISE+UKIDSS+DECaLS & \citet{2018Natur.553..473B} & 1 & 2500.0 \\
    \hline
  \end{tabular}\\
  \begin{minipage}{14.5cm}
   \textsuperscript{$c$}{Redshift range of the sample or, for small samples, approximate redshift range in which the survey is sensitive.}\\
  \textsuperscript{$b$}{Restricted to $M_{1450}>-26.73$ quasars to avoid overlap with the \citet{2016ApJ...829...33Y} sample.}\\
 \textsuperscript{$c$}{Used only in Section~\ref{sec:global} and Appendix~\ref{sec:conv} due to lack of spectroscopic redshifts for majority of the sample.}\\
 \textsuperscript{$d$}{Used only in Section~\ref{sec:global} due to roughly estimated selection function.}
 \end{minipage}
\end{table*}

@@

We started with compiling the samples of recent photometric rest-frame UV-optical
quasar surveys. The restriction to UV-optical surveys was mainly
driven by our science goal to characterise the UV luminosity function
of Type~1 quasars. X-ray-selected samples are less suited for this
purpose due to spectroscopic incompleteness and the $\sim 0.4$\,dex
scatter in the conversion from X-ray to UV luminosity
\citep{2010A&A...512A..34L, 2015MNRAS.453.1946G, 2016ApJ...819..154L}
that contributes significantly to the error budget in the UV
luminosity function of X-ray-selected samples unless rest-frame UV
photometry is incorporated \citep{2015AA...578A..83G}.  The individual
surveys and their main characteristics are listed in
Table~\ref{tab:samples}.  Figure~\ref{fig:qsos} presents a redshift
histogram of the contributing surveys.

@@

At $4.7\le z<5.5$ we combined several recent surveys, accounting for
sample overlap and updated selection functions. At the bright end of
the QLF we used the 99 quasars from the SDSS+WISE survey
\citep{2016ApJ...829...33Y} that have $M_{1450}<-26.73$ in our adopted
cosmology.  For these 99 quasars selected from 14,555\,deg$^2$ we
adopted the \citet{2016ApJ...829...33Y} selection function.  The
\citet{2016ApJ...829...33Y} sample partially overlaps with the SDSS
DR7 sample from \citet{2013ApJ...768..105M}, so to avoid
double-counting quasars we used the latter sample only at $M_{1450}\ge
-26.73$, yielding 104 additional $4.7\le z<5.5$ quasars selected in
6248\,deg$^2$. We used the $z\sim 5$ SDSS DR7 selection function from
\citet{2013ApJ...768..105M} that supersedes the one from
\citet{2006AJ....131.2766R} due to improved bandpass corrections and IGM
parameterization.  To these two bright-end samples we added the
faint-end sample from the \citet{2013ApJ...768..105M} SDSS Stripe~82
survey (62 uniformly selected $4.7\le z<5.5$ quasars in 235\,deg$^2$)
and two $4.7\le z<5.5$ quasars from \citet{2011ApJ...728L..26G},
adopting the respective selection functions.  We did not consider the
limit on the $z\sim 5$ QLF by \citet{2012ApJ...756..160I} due to
systematic errors in their selection
function\footnote{\citet{2012ApJ...756..160I} underestimated the
  dispersion in rest-frame UV quasar colours with respect to SDSS at
  all redshifts (their Figure~4). Contrary to their claim, photometric
  errors have a small effect on the colour distribution of SDSS
  quasars given the statistical errors of $<0.03$\,mag in $gri$ for 90
  per cent of the SDSS DR7 bright quasar sample ($i<19.1$) and a
  relative calibration error of $\sim 1$ per cent
  \citep{2008ApJ...674.1217P}.}.

@@

For a joint fit of the QLF it is necessary to homogenise the different
survey samples in absolute magnitude, and to convert their selection
functions to the same absolute magnitude system. For the analysis of
the quasar UV emissivity and to be consistent with published work at
$z>3$ we chose to convert all samples and selection functions to the
absolute AB magnitude at a rest frame wavelength $\lambda=1450$\,\AA
\begin{equation}\label{eq:absmag}
  M_{1450}\left(z\right) = m-5\log{\left(\frac{d_L\left(z\right)}
    {\mathrm{Mpc}}\right)}-25-K_{m,1450}\left(z\right),
\end{equation}
with the luminosity distance
\begin{equation}
  d_L(z)=(1+z)\frac{c}{H_0}\int_0^z\frac{\mathrm{d}z^\prime}
  {\sqrt{\Omega_\mathrm{m}(1+z^\prime)^3+\Omega_\Lambda}}
  \label{eqn:dl}
\end{equation}
to a quasar at redshift $z$, the apparent magnitude $m$ in a filter used
in the survey, and the bandpass correction $K_{m,1450}\left(z\right)$
\citep{1956AJ.....61...97H, 1968ApJ...154...21O, 2000A&A...353..861W,
  2002astro.ph.10394H}. For the bandpass correction we used a
combination of the \citet{2015MNRAS.449.4204L} stacked quasar spectrum at
$\lambda<2500$\,\AA, and the \citet{2001AJ....122..549V} quasar
composite spectrum at longer wavelengths to cover the lowest
redshifts. The samples from SDSS and BOSS are defined in the SDSS $i$
band, while 2SLAQ is defined in the $g$ band. At $z>4.7$ we adopted
the SDSS $z$ band magnitude (in the following denoted $z_\mathrm{AB}$)
for SDSS DR7 quasars to avoid additional corrections due to the
Ly$\alpha$ forest. Figure~\ref{fig:kcorr} shows our bandpass
corrections for SDSS, BOSS and 2SLAQ as a function of redshift. We
ignored the luminosity dependence of the bandpass correction due to
the known anticorrelation of emission line equivalent width and
luminosity \citep{1977ApJ...214..679B}.  While the
\citet{2015MNRAS.449.4204L} spectrum is for luminous ($M_{1450}\simeq
-27.2$) quasars, UV composite spectra including fainter quasars
\citep{2002ApJ...565..773T, 2012ApJ...752..162S, 2014ApJ...794...75S}
give similar values, such that our bandpass corrections remain
applicable at $M_{1450}\la -24$. Empirical luminosity-dependent
bandpass corrections show a $\la 0.2$\,mag variation over $\sim 4$
orders of absolute magnitude depending on redshift and the filter, and
with a $\sim 0.2$\,mag intrinsic scatter due to individual quasar-to-quasar
variations \citep{2013ApJ...773...14R, 2013ApJ...768..105M,
  2013A&A...551A..29P}.

@@

For 2SLAQ we corrected for magnitude-dependent spectroscopic coverage
in the two survey areas \citep[$f_\mathrm{c,NGP}\left(g\right)$ and
  $f_\mathrm{c,SGP}\left(g\right)$; Figure~4
  in][]{2009MNRAS.392...19C} and spectroscopic redshift success
\citep[$f_\mathrm{s,2SLAQ}\left(g\right)$; Figure~6b
  in][]{2009MNRAS.392...19C} by multiplying them with the photometric
selection function, resulting in two area-specific 2SLAQ selection
functions
$f_\mathrm{NGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,NGP}f_\mathrm{s,2SLAQ}$
and
$f_\mathrm{SGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,SGP}f_\mathrm{s,2SLAQ}$
that are relevant for the QLF.  The $z<4.7$ SDSS photometric selection
function was modified to include known imaging incompleteness to
$f_{\mathrm{SDSS},z<4.7}=0.95f_{\mathrm{p,SDSS},z<4.7}$
\citep{2006AJ....131.2766R}. The BOSS colour-selected sample contains
quasars with $f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}\ge 0.85$
\citep{2013ApJ...773...14R}, and we adopted
$f_\mathrm{BOSS}=\overline{f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}}f_\mathrm{p,BOSS}=0.962f_\mathrm{p,BOSS}$. \citet{2010ApJ...710.1498G}
presented two area-specific photometric selection functions due to
different filters employed, and more follow-up spectroscopy was
reported in \citet{2011ApJ...728L..26G}. We accounted for remaining
spectroscopic incompleteness at $R>23$, yielding the final selection
functions $f_\mathrm{NDWFS}$ and $f_\mathrm{DLS}$. The updated $z\sim
5$ SDSS photometric selection function \citep{2013ApJ...768..105M} was
modified to include imaging and spectroscopic incompleteness, yielding
$f_{\mathrm{SDSS},z\sim 5}=0.95^2f_{\mathrm{p,SDSS},z\sim 5}$. In the
deeper $z\sim 5$ SDSS Stripe~82 survey the spectroscopic
incompleteness is larger and magnitude-dependent \citep[Figure~14
  in][]{2013ApJ...768..105M}, resulting in $f_{\mathrm{S82},z\sim
  5}=0.95f_{\mathrm{s,S82},z\sim
  5}\left(i\right)f_{\mathrm{p,S82},z\sim 5}$. Likewise, imaging and
magnitude-dependent spectroscopic incompleteness was factored into the
\citet{2016ApJ...829...33Y} photometric selection function (their
Figures~5 and 7), resulting in
$f_\mathrm{SDSS+WISE}=0.97f_\mathrm{s,SDSS+WISE}\left(z_\mathrm{AB}\right)f_\mathrm{p,SDSS+WISE}$.
We obtained a rough estimate of the
\citet{2015AA...578A..83G} selection function by comparing the
corrected and observed QLFs, i.e.\ taking
$f_\mathrm{GOODS-S}=\phi_\mathrm{obs}/\phi_\mathrm{corr}$ (see their
Table~3).
Finally, for the three $z>6.5$ quasars we assumed a selection function of
unity in a range of $z$ and $M_{1450}$ estimated by the respective survey teams
(private communication).

@@

After homogenising the samples and selection functions we are in a position
to compute the UV luminosity function of AGN. We begin our analysis by
deriving luminosity functions in bins of redshift and absolute magnitude.
For simplicity we will use the notation $M\equiv M_{1450}$ in the following.

@@

In a magnitude bin $[M_\mathrm{min}, M_\mathrm{max})$, and redshift
bin $[z_\mathrm{min}, z_\mathrm{max})$, we define the luminosity
function as \citep{2000MNRAS.311..433P}
\begin{equation}
  \phi \equiv \frac{N}{V_\mathrm{bin}},
\end{equation}
where $N$ is the number of quasars with magnitude
$M_\mathrm{min}\leq M<M_\mathrm{max}$ and redshift
$z_\mathrm{min}\leq z<z_\mathrm{max}$, and
\begin{equation}
  V_\mathrm{bin} = \int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z},
  \label{eqn:vi}
\end{equation}
is the effective volume of the bin. The inclusion of the survey selection function
$f(M,z)$ (Section~\ref{sect:datahom}) in Equation~(\ref{eqn:vi}) accounts for
what are sometimes called ``incomplete bins''
\citep{2006AJ....131.2766R}.  The comoving volume element $\mathrm{d}V/\mathrm{d}z$ is
given by
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z}=\frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}\times A\times\frac{4\pi}{41253},
\end{equation}
where $A$ is the survey area in deg$^2$, and 
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}=\frac{c}{H_0}\frac{d_L^2\left(z\right)}
       {\left(1+z\right)^2\sqrt{\Omega_\mathrm{m}\left(1+z\right)^3+\Omega_\Lambda}}
  \label{eqn:dvdzdo}
\end{equation}
denotes the comoving volume element per unit solid angle
\citep{1999astro.ph..5116H}.

The luminosity function $\phi$ has units of
$\mathrm{cMpc}^{-3}\mathrm{mag}^{-1}$.  We evaluate the double
integral in Equation~(\ref{eqn:vi}) by the Euler method, i.e., by
simply summing over the ``tiles'' of the selection function map,
without interpolating between the redshift and luminosity values of
neighbouring tiles. 
%%GW: We should add that interpolation of selection functions is not straightforward (strong gradients) and not unique. Many selection functions are very coarse. V=0 means that the selection function has systematic errors (a QSO was found where the selection probability is zero). By removing them we avoid numerical issues, but underestimate the QLF. Surveys need to report selection functions that are consistent with their actual samples.
This may result in $V_i=0$ for some quasars, in
which case we remove them from our analysis.
%%GW: Poisson statistics will underestimate the error since uncertainties in the selection function are not considered (all surveys assume a fixed selection function). For small surveys the Poisson error dominates, but for large surveys the error is driven by systematics in the selection function. These facts have been neglected so far, and we should make this point very clearly.
In each bin, we estimate the uncertainty in the luminosity function by
assuming Poisson statistics \citep{1986ApJ...303..336G} for the number
of quasars, i.e.\ assuming negligible uncertainty in the selection function.
The resultant binned luminosity function estimates are shown by the
points in Figure~\ref{fig:mosaic}.

%%GW: I've sharpened the discussion below.
As seen in Figure~\ref{fig:mosaic}, the distribution of luminosity
function values in each redshift bin are suggestive of a double power
law form for the QLF.  We will fit such a form below. 
However, in several redshift bins we note a suspicious decline of the
luminosity function at the faint limit of several surveys, for example in
the 2SLAQ sample at $z<2.2$, and in the SDSS sample at $z<1.8$ and $z\sim 4$.
The inconsistency between the SDSS faint end and the deeper
2SLAQ QLF indicates that the SDSS selection function is
systematically overestimated at its magnitude limit.
At $z\sim 4$ the SDSS faint end QLF is inconsistent with the fainter
\citet{2011ApJ...728L..26G} QLF. We identify such discrepant bins
by eye and discard them from our analysis.
%%GW: We should also set the selection functions to zero in these bins.
The discarded magnitude bins are shown in Figure~\ref{fig:mosaic} by open circles.

%%GW: 2.5<z<2.6 QSOs are missing in this plot! Also, there are no open symbols for BOSS, but the numbers indicate that you excluded some in several redshift bins. The total numbers per redshift bin are not matching the number of BOSS QSOs.
\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    % mosaic.py 
  \end{center}
  \caption{Homogenised quasar luminosity functions at $\lambda=1450$\,\AA\ in redshift bins from
    $z=0.1$ to $6.5$.  The symbols show our inferred binned
    luminosity functions from various data sets:
    \citet[red]{2010AJ....139.2360S},
    \citet[green]{2009MNRAS.392...19C},
    \citet[dark blue]{2013ApJ...773...14R},
    \citet[light blue]{2011ApJ...728L..26G},
    \citet[yellow]{2016ApJ...829...33Y},
    \citet[brown]{2013ApJ...768..105M},
    \citet[pink]{2016ApJ...833..222J},
    \citet[orange]{2010AJ....139..906W}, and
    \citet[grey]{2015ApJ...798...28K}.  Open circles in
    corresponding colours indicate excluded incomplete magnitude bins for
    the respective data sets. In each panel we give the number
    of considered quasars in the respective redshift bin,
    with the number in parentheses indicating the total number of quasars before
    excluding incomplete magnitude bins.
    In each redshift bin, the black
    curve shows the best-fit double power law model, which is
    represented by the median of the posterior probability
    distribution function.  The grey shaded area shows the
    one-sigma (68.26\%) uncertainty. 
    See Sections~\ref{sec:binnedlf} and \ref{sec:bins} for further
    details.}
  \label{fig:mosaic}
\end{figure*}
